Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-09-04 jso.qmd
add practical guide for computing smoothness and squintability

@@

```{r}
#| eval: false
#| echo: false
jso_optima <- save_history(flea[,1:6], 
                           guided_tour(holes()), 
                           search_f = search_jellyfish)
```
<!--
This section discusses the insights in implementation yielded from the previous sections to guide the usage of projection pursuit guided tour amongst practitioners. Smoothness and squintability in a projection pursuit problem can now be measured given the target shape, the chosen index function, and the data dimension. In practice, these metrics can be used to characterise the difficulty of the optimisation task and inform the choice of optimiser. For instance, a large squintability value indicates a distinct difference in index values between optimal regions and others. When one jellyfish enters the optimal region, the high index value it generates will be clearly distinguishable from spurious values due to this pronounced difference and lead other jellyfish to move toward the optimal region. In contrast, a smaller squintability value results in a less distinct separation between desirable and undesirable regions, making it more difficult for other jellyfish to accurately navigate toward the target area. In addition, a larger smoothness value implies a smoother index function, which suggests the use of both gradient-based and random sampling optimisers.

>>>>

The `ferrn` package provides functionality for computing the smoothness and squintability metrics. Both cases require sampling random bases using the `sample_bases()` function, followed by  calculating the specific metric with either `calc_smoothness()` or `calc_squintability()`.

### Smoothness

To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance). The following code samples 300 bases for using the `holes` index to find the pipe shape in the 6D `sine1000` dataset:

```{r}
#| echo: true
library(GpGp)
library(fields)
library(tourr)
library(ferrn)
basis_smoothness <- sample_bases(
  idx = "holes", data = sine1000, n_basis = 300, 
  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6))
basis_smoothness |> head(3)
```

The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value). To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.

The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric. The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`. The following code calculates the smoothness metric from the previously sampled bases:

```{r}
#| echo: true
(sm_res <- calc_smoothness(basis_smoothness))
```

The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.

### Squintability

Sampling bases for calculating squintability includes an additional step of interpolating between the sampled bases and the best basis. This is specified through setting the `step_size` and `min_proj_dist` arguments in the `sample_bases()` function to non-NA numerical values. Since the projection distance typically ranges from 0 to 2, it is recommended to set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5 to ensure a meaningful interpolation length.

The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`). A `bin_width` argument is required to average the index values over the projection distance before the fit. The following code samples 100 bases for calculating squintability in the same scenario:

```{r}
#| echo: true
basis_squint <- sample_bases(
   idx = "holes", data = sine1000, n_basis = 100, 
  best = matrix(c(0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 1), nrow = 6),
  step_size = 0.1, min_proj_dist = 1.5)
calc_squintability(basis_squint, method = "nls", bin_width = 0.01)
```

The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case. For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.

@@

```{r}
#| eval: false
#| echo: false
jso_optima <- save_history(flea[,1:6], 
                           guided_tour(holes()), 
                           search_f = search_jellyfish)
```

<!--
This section discusses the insights in implementation yielded from the previous sections to guide the usage of projection pursuit guided tour amongst practitioners. Smoothness and squintability in a projection pursuit problem can now be measured given the target shape, the chosen index function, and the data dimension. In practice, these metrics can be used to characterise the difficulty of the optimisation task and inform the choice of optimiser. For instance, a large squintability value indicates a distinct difference in index values between optimal regions and others. When one jellyfish enters the optimal region, the high index value it generates will be clearly distinguishable from spurious values due to this pronounced difference and lead other jellyfish to move toward the optimal region. In contrast, a smaller squintability value results in a less distinct separation between desirable and undesirable regions, making it more difficult for other jellyfish to accurately navigate toward the target area. In addition, a larger smoothness value implies a smoother index function, which suggests the use of both gradient-based and random sampling optimisers.

### 2024-09-04 jso.tex
add practical guide for computing smoothness and squintability

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{orcidlink}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{assumption}{\protect\assumptionname}
\newtheorem{claim}{\protect\claimname}
\newtheorem{condition}{\protect\conditionname}
\newtheorem{conjecture}{\protect\conjecturename}
\newtheorem{cor}{\protect\corollaryname}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{example}{\protect\examplename}
\newtheorem{lem}{\protect\lemmaname}
\newtheorem{notation}{\protect\notationname}
\newtheorem{problem}{\protect\problemname}
\newtheorem{prop}{\protect\propositionname}
\newtheorem{rem}{\protect\remarkname}
\newtheorem{thm}{\protect\theoremname}
\providecommand{\assumptionname}{Assumption}
\providecommand{\claimname}{Claim}
\providecommand{\conditionname}{Condition}
\providecommand{\conjecturename}{Conjecture}
\providecommand{\corollaryname}{Corollary}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\notationname}{Notation}
\providecommand{\problemname}{Problem}
\providecommand{\propositionname}{Proposition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\begin{tcolorbox}[enhanced jigsaw, toprule=.15mm, coltitle=black, bottomrule=.15mm, titlerule=0mm, colbacktitle=quarto-callout-note-color!10!white, rightrule=.15mm, leftrule=.75mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, breakable, title={Algorithm: Jellyfish Optimizer Pseudo Code}, left=2mm, opacitybacktitle=0.6, opacityback=0, bottomtitle=1mm, arc=.35mm, colback=white]

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{orcidlink}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{assumption}{\protect\assumptionname}
\newtheorem{claim}{\protect\claimname}
\newtheorem{condition}{\protect\conditionname}
\newtheorem{conjecture}{\protect\conjecturename}
\newtheorem{cor}{\protect\corollaryname}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{example}{\protect\examplename}
\newtheorem{lem}{\protect\lemmaname}
\newtheorem{notation}{\protect\notationname}
\newtheorem{problem}{\protect\problemname}
\newtheorem{prop}{\protect\propositionname}
\newtheorem{rem}{\protect\remarkname}
\newtheorem{thm}{\protect\theoremname}
\providecommand{\assumptionname}{Assumption}
\providecommand{\claimname}{Claim}
\providecommand{\conditionname}{Condition}
\providecommand{\conjecturename}{Conjecture}
\providecommand{\corollaryname}{Corollary}
\providecommand{\definitionname}{Definition}
\providecommand{\examplename}{Example}
\providecommand{\lemmaname}{Lemma}
\providecommand{\notationname}{Notation}
\providecommand{\problemname}{Problem}
\providecommand{\propositionname}{Proposition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\begin{tcolorbox}[enhanced jigsaw, coltitle=black, rightrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, breakable, opacitybacktitle=0.6, arc=.35mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, opacityback=0, bottomtitle=1mm, titlerule=0mm, toptitle=1mm, bottomrule=.15mm, toprule=.15mm]

@@

The \texttt{ferrn} package provides functionality for computing the
smoothness and squintability metrics. Both cases require sampling random
bases using the \texttt{sample\_bases()} function, followed by
calculating the specific metric with either \texttt{calc\_smoothness()}
or \texttt{calc\_squintability()}.

\subsubsection{Smoothness}\label{smoothness}

To sample bases for calculating smoothness, the required arguments
include \texttt{idx} (the index function used), \texttt{data} (the
dataset), \texttt{n\_basis} (the number of random bases to sample), and
\texttt{best} (the best basis to compute projection distance). The
following code samples 300 bases for using the \texttt{holes} index to
find the pipe shape in the 6D \texttt{sine1000} dataset:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(GpGp)}
\FunctionTok{library}\NormalTok{(fields)}
\FunctionTok{library}\NormalTok{(tourr)}
\FunctionTok{library}\NormalTok{(ferrn)}
\NormalTok{basis\_smoothness }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
  \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{300}\NormalTok{, }
  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{))}
\NormalTok{basis\_smoothness }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# PP index:                holes
# No. of bases:            3
# interpolation step size: NA
# Min. proj. dist.:        NA
     id basis          dist index
  <int> <list>        <dbl> <dbl>
1     1 <dbl [6 x 2]>  1.15 0.676
2     2 <dbl [6 x 2]>  1.77 0.790
3     3 <dbl [6 x 2]>  1.79 0.778
\end{verbatim}

The output of \texttt{sample\_bases()} is a data frame with four
columns: \texttt{id} (the basis ID), \texttt{basis} (the basis matrix
sampled), \texttt{dist} (the projection distance to the best basis), and
\texttt{index} (the index value). To speed up the computation of index
values, parallellasation is available through the
\texttt{parallel\ =\ TRUE} argument.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the projection distances to compute the smoothness
metric. The starting parameter for the fit can be specified using the
\texttt{start\_params} argument, and additional Gaussian process
arguments can be provided through \texttt{other\_gp\_params}. The
following code calculates the smoothness metric from the previously
sampled bases:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(sm\_res }\OtherTok{\textless{}{-}} \FunctionTok{calc\_smoothness}\NormalTok{(basis\_smoothness))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# PP index:     holes
# No. of bases: 300 [6 x 2]
    variance range smoothness nugget convergence
       <dbl> <dbl>      <dbl>  <dbl> <lgl>      
1 0.00000672  18.1       1.03  1138. TRUE       
\end{verbatim}

The smoothness metric is the smoothness parameter from the Gussian
process and all details of the fit can be accessed using
\texttt{attr(sm\_res,\ "fit\_res")}.

\subsubsection{Squintability}\label{squintability}

Sampling bases for calculating squintability includes an additional step
of interpolating between the sampled bases and the best basis. This is
specified through setting the \texttt{step\_size} and
\texttt{min\_proj\_dist} arguments in the \texttt{sample\_bases()}
function to non-NA numerical values. Since the projection distance
typically ranges from 0 to 2, it is recommended to set
\texttt{step\_size} to 0.1 or lower, and \texttt{min\_proj\_dist} to be
at least 0.5 to ensure a meaningful interpolation length.

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fit. The following code samples
100 bases for calculating squintability in the same scenario:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{basis\_squint }\OtherTok{\textless{}{-}} \FunctionTok{sample\_bases}\NormalTok{(}
   \AttributeTok{idx =} \StringTok{"holes"}\NormalTok{, }\AttributeTok{data =}\NormalTok{ sine1000, }\AttributeTok{n\_basis =} \DecValTok{100}\NormalTok{, }
  \AttributeTok{best =} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{nrow =} \DecValTok{6}\NormalTok{),}
  \AttributeTok{step\_size =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{min\_proj\_dist =} \FloatTok{1.5}\NormalTok{)}
\FunctionTok{calc\_squintability}\NormalTok{(basis\_squint, }\AttributeTok{method =} \StringTok{"nls"}\NormalTok{, }\AttributeTok{bin\_width =} \FloatTok{0.01}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# PP index:     holes
# No. of bases: 100 -> 1853
# method:       nls
  theta1 theta2 theta3 theta4 squint
   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
1  0.274  0.960   2.21  0.612  0.812
\end{verbatim}

The output provides the estimated parameters for the scaled sigmoid
function (\(\theta_1\) to \(\theta_4\)) and the squintability metric for
the parametric case. For the non-parametric case, it shows the
(\texttt{max\_d}), the corresponding projection distance
(\texttt{max\_dist}), and teh squintability metric as their products.
